Cognitive and ecological aspects determine cleaner performance

### 2021-10-29 manuscript_1.0.Rmd
More or  less done with the methods. Trying out to adapt the model for using BayesTools

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(here)
```

@@

In order to fit the model parameters to the data we used Markov Chain Monte Carlo (MCMC) with 
a Metropolis sampler to estimate the posterior distribution of the model parameters. We used the data from @triki_Biological_2019 to fit the parameters of the  model analyzed in @quinones_Reinforcement_2019. The data consist a of the number of times that set of cleaners chose to
feed of a client plate in the context of the market task. This set of cleaners were collected under different locations and in each one of those locations environmental variables were measured. These environmental variables recorded from each location are used as input for the computational model. Given these inputs, the computational model produces a time series of choices in the market task. In the model the preference for the two option changes as the algorithm learns based on the reward that the two options provide. Thus, we run the model for 10000 steps and use the las 3000 steps as the preference developed in the model for visitors given the environmental variables. We use this preference as the predicted probability of choosing a visitor in the data, and use it to calculate the likelihood that the number of visitor choices obtained from each cleaner is generated by the model. We sum over all the cleaners the loglikehoods and use the total for the acceptance of new parameters in the MCMC.

The model consist of a set individual-simulations where individuals face a series of choices between two options. Individuals in the model represent cleaners, and the options they face simulate the natural conditions of a cleaning market. Thus, individuals in the model choose to prioritize  different clients. These clients are classified as resident and visitors. Residents are those clients that stay in the cleaning queue when they are not given priority; while visitors leave the queue (with a certain probability) after they are not given priority. Individuals obtain reward from cleaning a client regardless of the type.  Everytime individuals face and make a choice they update the probability of making that same choice. The update is based on the difference between the expectation of reward and the actual reward the individual obtains (prediction error); and it is carried in the direction that would lead to more reward given the new information. Further details of the model implementation can be found in the paper [@quinones_Reinforcement_2019]. The model analysis points to the relevance of two parameters. First, $\gamma$ measures how much individuals include future rewards in their desicion updates. If $\gamma=0$, individuals only use the immediate reward obtained from a cleaning interaction. As $\gamma$ increases individuals include more the reward obtained from the subsequent choices. That amounts to calculating and using for deision making future expected rewards. Second, $\eta$ measures how much individuals include in their reward the fleeing behaviour of visitor clients as a negative component. Both of these components allow individuals to use in ther estimates the future effects of their choices. The model analysis shows that without either of them, learning individuals cannot develop a preference for the visitor client. The analysis contrast the effect of these two parameters agaisnt variation in the relative abundance of the two client types. The analysis shows that under very high client abundance the presence of  future rewards ($\gamma>0$) no longer favours the preference for the ephemeral clients, this effect does not occur with negative reward mediated by client behaviour ($\eta$). Thus, the analysis suggests that future reward as an explanation for cleaner ability to prefer visitors is supported by @triki_Biological_2019 data. However, a combination of future and negative rewards, could still be a better explanation for the data. In the current analysis, we will use the data from @triki_Biological_2019 to fit some of the model parameters, particularly relevant we will fit $\gamma$ and $\eta$ to find out whether each or a combination of these effects is a better explanation for the pattern seen in the data.

@@

In order to capture with the model the relationship between the characteristics of the market (environment) and the cleaner's performance we need to translate the absolute abundances of the data to relative abundances. This is because in the model, relative abundances of clients define not only the probability of residents and visitors, but also how often the cleaning station is empty (e.g. there are no clients to be cleaned). The more cleaners, the more options available for the clients and hence higher probability of empty cleaning stations. How often the cleaning stations are empty in turns influences the value that cleaners can expect from their choices. However, the frequency of empty spots in the cleaning station does not only depend on cleaner relative abundances but also on the frequency with which clients visit the station. The client visitation frequency will likely vary among different clients depnding on parasite abundances and loads. There are variables for which we do not have field estimates. Thus, we use
computed a measure of relative cleaner abundance for each site relative to absolute abundances and multiplied by scalling constant that changes the range of the variable. We fit the value os the scalling constant as part of the statistical inference. As for the client abundances for residents and visitors we computed a relative measure with respect to the total client abundance, and weigthted that by the rescaled relative absence of cleaners (1-relative cleaner abundance). Thus, all three measures of relative abundance sum up to 1, and can be used in the model as a proxy for the probability of having different options in the cleaning station.

@@

```{r, fig1}
knitr::include_graphics(here("Simulations","ABCclean_gam_nRew_sca_",                            "round_gamma_0.3regRew_0.11scaC_133.5alphA_0.05alphC_0.05_contour_ggplot.png"))
```

>>>>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(here)
```

@@

In order to fit the model parameters to the data we used Markov Chain Monte Carlo (MCMC) with 
a Metropolis sampler to estimate the posterior distribution of the model parameters. We used the data from @triki_Biological_2019 to fit the parameters of the  model analyzed in @quinones_Reinforcement_2019. The data consist a of the number of times that a set of cleaners chose to
feed of a visitor client plate in the context of the market task. This set of cleaners were collected in different locations and in each one of those locations environmental variables were measured. These environmental variables recorded from each location are used as input for the computational model. Given these inputs, the computational model produces a time series of choices in the market task. In the model the preference for the two option changes as the algorithm learns based on the reward that the two options provide. Thus, we run the model for 10000 steps and use the las 3000 steps as the preference developed in the model for visitors given the environmental variables. We use this preference as the predicted probability of choosing a visitor in the data, and use it to calculate the likelihood that the number of visitor choices obtained from each cleaner is generated by the model. We sum over all the cleaners the loglikehoods and use the total for the acceptance of new parameters in the MCMC.

The model consist of a set individual-simulations where individuals face a series of choices between two options. Individuals in the model represent cleaners, and the options they face simulate the natural conditions of a cleaning market. Thus, individuals in the model choose to prioritize  different clients. These clients are classified as resident and visitors. Residents are those clients that stay in the cleaning queue when they are not given priority; while visitors leave the queue (with a certain probability) after they are not given priority. Individuals obtain reward from cleaning a client regardless of the type.  Everytime individuals face and make a choice they update the probability of making that same choice. The update is based on the difference between the expectation of reward and the actual reward the individual obtains (prediction error); and it is carried in the direction that would lead to more reward given the new information. Further details of the model implementation can be found in the paper [@quinones_Reinforcement_2019]. The model analysis points to the relevance of two parameters. First, $\gamma$ measures how much individuals include future rewards in their desicion updates. If $\gamma=0$, individuals only use the immediate reward obtained from a cleaning interaction. As $\gamma$ increases individuals include more the reward obtained from the subsequent choices. That amounts to calculating and using for decision making the future expected rewards. Second, $\eta$ measures how much individuals include in their reward the fleeing behaviour of visitor clients as a negative component. Both of these parameters allow individuals to use in ther estimates the future effects of their choices. The model analysis shows that without either of them, learning individuals cannot develop a preference for the visitor client. The analysis contrast the effect of these two parameters against variation in the relative abundance of the two client types. The analysis shows that under very high client abundance the presence of  future rewards ($\gamma>0$) no longer favours the preference for the ephemeral clients, this effect does not occur with negative reward mediated by client behaviour ($\eta$). Thus, the analysis suggests that future reward as an explanation for cleaner ability to prefer visitors is supported by @triki_Biological_2019 data. However, a combination of future and negative rewards, could still be a better explanation for the data. In the current analysis, we will use the data from @triki_Biological_2019 to fit some of the model parameters, particularly relevant we will fit $\gamma$ and $\eta$ to find out whether each or a combination of these effects is a better explanation for the pattern seen in the data.

@@

In order to capture with the model the relationship between the characteristics of the market (environment) and the cleaner's performance we need to translate the absolute abundances of the data to relative abundances. This is because in the model, relative abundances of clients define not only the probability of residents and visitors, but also how often the cleaning station is empty (e.g. there are no clients to be cleaned). The more cleaners, the more options available for the clients and hence higher probability of empty cleaning stations. How often the cleaning stations are empty in turns influences the value that cleaners can expect from their choices. The more empty the empty cleaning stations are, the more it is to choose a client that will not stay.  However, the frequency of empty spots in the cleaning station does not only depend on cleaner relative abundances but also on the frequency with which clients visit the station. The client visitation frequency will likely vary among different clients depnding on parasite abundances and loads. These are variables for which we do not have field estimates. Thus, we
computed a measure of relative cleaner abundance for each site relative to absolute abundances and multiplied by a scalling constant that changes the range of the variable. We fit the value os the scalling constant as part of the statistical inference. As for the client abundances for residents and visitors we computed a relative measure with respect to the total client abundance, and weigthted that by the rescaled relative absence of cleaners (1-relative cleaner abundance). Thus, all three measures of relative abundance sum up to 1, and can be used in the model as a proxy for the probability of having different options in the cleaning station.

@@

Posterior distribution samples taken from the MCMC (Fig.\ref{fig:estimValues}) show data from the market taks experiments fit best with a learning model that includes the use of both future reward ($\gamma>0$) and negative reward triggered by visitor clients leaving the station ($\eta>0$). The value of negative reward is, however, small.

```{r estimValues,fig.cap="Estimates posterior distributions for parameter values of $gamma$, $eta$ and the scalling constant. We show the kernet density estimates, the mode (black line) and the 95% credible interval for the three paramteres"}
scenario<-"ABCclean_gam_nRew_sca_2_"
file<-list.files(here("Simulations",scenario),full.names = TRUE,
                      pattern = "MCMC.png")
knitr::include_graphics(file)
```

```{r predictions, fig.cap="Observed and predicted probability of choosing a visitor. Left-hand side pane: color contour shows the prediction of the learning model using the mode of the posterior distributions of parameters recovered by the statistical analysis. Dots show the frequency of visitor choices for the 12 locations, as well as the corresponding relative cleaner abundance (x axis) and frequency of visitors leaving the station (y axis). Right-hand side panel: scatter plot of the observed and predicted probabilities of choosing a visitor for 12 locations. Line correspond to a perfect match between observed and predicted probabilities."}
scenario<-"ABCclean_gam_nRew_sca_2_"
file<-list.files(here("Simulations",scenario),full.names = TRUE,
                      pattern = ".png")
knitr::include_graphics(file[3])
```
